---
title: "Dramatic transitions"
output: html_notebook
author: baj
---

# Our questions for the plays of Holberg

To examine dramatic transitions in Holberg's plays, we will try to find :
the first line in each scene
the last line in each scene

# From TEI to JSONL

As we described in the notebook "How to read plays into R", we used the XLS conversion to convert our files in TEI-format to JSONL-format, since this format is more suitable for R.

# Read a play into R

Load the play from the folder into JSONL-format.
Load the JSONL data format

```{r}
library(tidyverse)
library(ndjson)
library(here)

source(here("src", "p017-functions.R"))

my_dir <-"test-data"
my_play <- "Plutus_mod.jsonl"
read_play_jsonl(here(my_dir, my_play)) -> play
play # Call "play" to see the table that has now been assigned to that variable.

```

## Find the first sentence in each scene.

```{r}
  # Create a new variable to store first sentences in scenes, that we will now extract.
first_sentence <- play %>%
  # Use group_by to examine the play scene by scene.
 group_by(act_number, scene_number) %>%
  # Filter out all lines that are not character-lines (such as stage directions).
 filter(!is.na(spoke)) %>%
  # To get all lines in each scene in order, we sort our data by each group (scene).
 arrange(index, .by_group = TRUE) %>%
  # Create a new index in each group that reflects the right order, each starting with the first line of each scene.
 mutate(group_row_num = row_number()) %>%
  # Keep only the first line in each group, so that we have only the first line of all of our scenes.
 filter(group_row_num == 1) %>%
  # Lastly, remove the index again - it is no longer needed.
 select(- group_row_num )

first_sentence
```

We have to save the result. 
Also, when saving it, remember to change the name of the file so we do not overwrite the play itself.

```{r}
write.csv(first_sentence, file = here("graphs/Første og sidste sætning sagt/first_sentence_Plutus.csv"))
```

Surely, we can extract the last sentences of all scenes in the same way?

```{r}
  # Create a new variable to store last sentences, which we will now extract.
last_sentence <- play %>%
  # Use group_by to examine the play scene by scene.
 group_by(act_number, scene_number) %>%
  # Filter out all lines that are not character-lines (such as stage directions).
 filter(!is.na(spoke)) %>%
  # Again, we sort our data by group, but now we want it in the reverse order, so the last sentences of each scene to come first.
 arrange(desc(index), .by_group = TRUE) %>%
  # Again, create a new index in each group that reflects the right order, each now starting with the last line of each scene.
 mutate(group_row_num = row_number()) %>%
  # Keep only the first line in each group, so that we have only the last line of all of our scenes.
 filter(group_row_num == 1) %>%
  # Lastly, remove the index again - it is no longer needed.
 select(- group_row_num )

write.csv(last_sentence, file = here("graphs/Første og sidste sætning sagt/last_sentence_Plutus.csv"))
```





