---
title: "Speakers"
output: html_notebook
author: baj
---

# Networks
We want to make visualizations of networks, where all speakers are divided by title, act, and scene.

# Read one play and all plays into R
We may find characteristics of just one play or of all plays together with the following lines of code.
First, we import the required libraries and load files in JSONL-format.


```{r}
library(tidyverse)
library(ndjson)
library(here)
library(fs)
library(tidytext)
library(xslt)
library(readxl)

barselstuen_raw <- ndjson::stream_in(here("test-data/Barselstuen_mod.jsonl")) %>% tibble()

source(here("src", "p017-functions.R"))

read_plays_jsonl (here("test-data")) ->allplays
```

# Looking at the data

```{r}
 barselstuen_raw
```

We may save the JSONL file as a csv in the following way:

```{r}
 write.csv(barselstuen_raw, file = here("csv/barselstuen_speakers.csv"))

```

## Finding the names of all characters with lines

Let's count all the named characters with lines in the plays.
We can make a list of speakers, where the sum of speakers will be the length of the list - although some characters may appear in the list more than once.
We may use the dictionary to remove the duplicates.

We can also make a table of speakers divided into acts and scenes. 

```{r}
here()

all_speakers <- allplays %>% 
  count(speaker)
all_speakers
# Get column names
colnames(all_speakers)
write.csv(all_speakers, file = here("csv/allspeakers.csv"))

# The below piece of code is only used when we examine one piece at a time.
speakers_by_scene <- allplays %>% 
  count(act, scene, speaker)
speakers_by_scene
write.csv(speakers_by_scene, file = here("csv/speakers_by_scene.csv"))
```

Next we load and run the excel sheet of characters in the plays:

```{r}
(variants <- read_excel(here("Rolleliste.xlsx")) %>% 
  unnest_tokens(variant, Alias, token = "regex", pattern = ", ") %>% 
    mutate(
      Karakter = tolower(Karakter),
      variant = tolower(variant)
    ))
# Merge the variations of names to one name (see "use-excel-for-name-variants.Rmd")
# Remember to use the file-name
(allplays %>% 
  mutate(speaker = tolower(speaker)) %>% 
  left_join(variants, by = c("speaker"="variant")) %>% 
  mutate(speaker = if_else(!is.na(Karakter), Karakter, speaker)) %>%
      filter(!is.na(speaker), !(speaker=="")) %>%
distinct -> allplays)

all_speakers <- allplays %>% 
  count(speaker)
all_speakers
# Get column names
colnames(all_speakers)
write.csv(all_speakers, file = here("csv/allspeakers uden navnevarianter.csv"))

```

## Find all plays that the characters are in

Let's now add a column where we count the number of plays that each character plays a part in.


```{r}
all_speakers_with_plays <- allplays %>% 
  count(speaker, docTitle, year) %>%
  group_by(speaker) %>%
  mutate(count_speaker = sum(n)) %>%
  count(docTitle, count_speaker) %>%
  summarise(plays = toString(unique(docTitle)), count_speaker, count_plays = sum(n)) %>%
  unique()
all_speakers_with_plays

write.csv(all_speakers_with_plays, file = here("csv/allspeakers uden navnevarianter med stykker.csv"))

```

## Find the names of all characters who are only mentioned in the stage directions

We want to count the number characters that are only mentioned in stage directions. Let's start by finding all lines of stage-directions. They are preceded by <stage> and by <speaker_stage>.

```{r}
# Find all lines preceded by <stage> and save them in the variable "all_stage_instructions"
# Find <stage>:
all_stage_instructions <- barselstuen_raw %>%
  select(stage) %>% filter(!is.na(stage))
all_stage_instructions
# write.csv(all_stage_instructions, file = here("output/barselstuen_all_stage_instructions.csv"))

# Find all lines preceded by <speaker_stage>, again, save them in the variable "all_stage_instructions"
# Find <speaker-stage>:
all_speaker_stage_instructions <- barselstuen_raw %>%
  select(speaker_stage) %>% filter(!is.na(speaker_stage))
all_speaker_stage_instructions

# Combine the two vectors that we have just extracted
combined_stage_vector <- c(as.vector(all_stage_instructions[['stage']]), as.vector(all_speaker_stage_instructions[['speaker_stage']]))
combined_stage_vector
write.csv(combined_stage_vector, here("output/barselstuen_combined_stage_instructions.csv"))

# Let's do the same, but this time we note from which scene in the play each stage direction is taken. 
# Find all lines preceded by <stage> by <scene> and save them in the variable "stage_instructions_by_scene"
# Find <stage> by <scene>:
stage_instructions_by_scene <- barselstuen_raw %>%
  select(act, scene, stage) %>% filter(!is.na(stage))
stage_instructions_by_scene
write.csv(stage_instructions_by_scene, file = here("output/barselstuen_stage_instructions_by_scene.csv"))

# Find all lines preceded by <speaker_stage> by <scene> and save them in the variable "stage_instructions_by_scene"
# Find <speaker-stage> by <scene>:
speaker_stage_instructions_by_scene <- barselstuen_raw %>%
  select(act, scene, speaker_stage) %>% filter(!is.na(speaker_stage))
speaker_stage_instructions_by_scene
write.csv(speaker_stage_instructions_by_scene, file = here("output/barselstuen_speaker_stage_instructions_by_scene.csv"))

```

So, there are 131 + 54 = 185 lines of stage directions in this play. Yet we also want to know the names of the characters mentioned in these. 

We can do that by searching for specific names - but what names do we search for? First, we can search for those characters that we found in the lines of stage directions preceded by <speaker>. However, we would miss some of those that are present, but do not speak.

