---
title: "Examples of visualization"
output: html_notebook
author: pmd
---

Import the necessary libraries:

```{r, message=FALSE}
library(tidyverse)
library(here)
library(ndjson)
library(xslt)
library(fs)
library(ggplot2)
library(readxl)
library(tidytext)

require(readr)  # For read_csv()
require(purrr)  # For map(), reduce()

source(here("src/p017-functions.R"))
```

# Read in the data

Read in the data of plays converted to JSONL

```{r}
read_plays_jsonl(here("test-data")) %>% select(filename, everything()) -> plays
```

## Read in the excel sheet of characters and variations of character-names

```{r}
variants <- read_excel(here("Rolleliste.xlsx")) %>% 
     unnest_tokens(variant, Alias, token = "regex", pattern = ", ") %>% 
    mutate(
      Karakter = tolower(Karakter),
      variant = tolower(variant)
    )
```

Combine the plays and the excel sheet:

```{r}
plays %>% 
  mutate(speaker = tolower(speaker)) %>% 
  left_join(variants, by = c("filename"="Filnavn", "speaker"="variant")) %>% 
  mutate(speaker = if_else(!is.na(Karakter), Karakter, speaker)) %>%
  distinct -> plays
```

# Visualization

Basically, we need to implement a function that does the following: `character_comparison_plot(<list of character names>, <list of character names>)`

```{r}
# Count words in "spoke"
plays %>%
  # Remove rows that are not dialogue and empty speakers
  filter(!is.na(spoke), speaker != "") %>%
  
  # Select only the columns, that we are interested in
  select(speaker, title, year, spoke) %>%
  
  # Add the number of spoken words
  mutate(n_spoken_words = str_count(spoke, '\\w+')) %>%
  # Remove the spoken words
  select(-spoke) %>%
  
  # Group by title and speaker
  group_by(title, year, speaker) %>%
  
  # Sum the words spoken by each speaker in each play
  # Lastly store the new data frame in a new reference for later use
    summarise(words = sum(n_spoken_words), title, year) %>%
    ##    ungroup() %>%
  distinct() -> summary

summary %>%
    ungroup(speaker) %>%
    summarise(words_play = sum(words)) -> totals

summary %>%
  full_join(totals) %>%
    summarise(freq = words / words_play, speaker, title, year, words, words_play) %>%
    ungroup() -> summary

summary

```

## Comparison of individual characters

```{r test}
  summary %>% 
  #filter(speaker %in% c("henrik","pernille","troels")) %>%
  filter(speaker %in%   c("henrik", "jeronimus", "leander")) -> tmp
  #arrange(year) %>%

  # Below follow two possible ways of visualizing the comparison.

  # Add year to title
  tmp %>% mutate(title = paste(year," ",title)) %>%
  ggplot(aes(fill = speaker, y = freq, x = title)) +
    geom_bar(stat="identity") +
    theme(legend.position="bottom") +
    labs(
      title = "Distribution of spoken words between few characters that reoccur across Holberg's oeuvre by year of publication",
      caption = "Source: DSL"
    ) +
  xlab("Year") +
  ylab("Pecentage of spoken words") +
  theme(axis.text.x=element_text(angle = -60, hjust = 0))

      ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_v1.pdf"), width=16, height=9)

  # We might also use year as a facet
  tmp %>%
  ggplot(aes(fill = speaker, y = freq, x = title)) +
    geom_bar(stat="identity") +
    theme(legend.position="bottom") +
    labs(
      title = "Distribution of spoken words between few characters that reoccur across Holberg's oeuvre by year of publication",
      caption = "Source: DSL"
    ) +
  xlab("Year") +
  ylab("Percentage of spoken words") +
  facet_wrap(~year, scales = 'free_x', ncol=nrow(summary), switch="x") +
  theme(axis.text.x=element_text(angle = -60, hjust = 0))

    ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_v2.pdf"), width=16, height=9)

```

## Comparison between groups of characters

```{r}
group_1 <- c("henrik", "jeronimus", "leander")
group_2 <- c("pernille", "leonora", "magdelone")

summary %>% 
  filter(speaker %in% c(group_1, group_2)) %>% 
  mutate(gruppe = if_else(speaker %in% group_1, "Henrik, Jeronimus og Leander","Pernille, Leonora og Magdelone")) -> tmp2
  #filter(speaker %in% c("henrik","pernille","leonora")) %>%
  #arrange(year) %>%
  
  # Below follow two possible ways of visualizing the comparison.

  # Add year to title
  tmp2 %>% mutate(title = paste(year," ",title)) %>%
    ggplot(aes(fill = gruppe, y = freq*100, x = title)) +
    geom_bar(stat="identity") +
    theme(legend.position="bottom") +
    labs(
      title = "Distribution of spoken words between few characters that reoccur across Holberg's oeuvre by year of publication",
      caption = "Source: DSL"
    ) +
  xlab("Year") +
  ylab("Percentage of spoken words") +
  labs(fill='Maskpair') +
  theme(axis.text.x=element_text(angle = -60, hjust = 0))
        ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_grupper_v1.pdf"), width=16, height=9)

  # We migt also use year as a facet
  tmp2 %>%
    ggplot(aes(fill = gruppe, y = freq*100, x = title)) +
    geom_bar(stat="identity") +
    theme(legend.position="bottom") +
    labs(
      title = "Distribution of spoken words between few characters that reoccur across Holberg's oeuvre by year of publication",
      caption = "Source: DSL"
    ) +
  xlab("Year") +
  ylab("Percentage of spoken words") +
  labs(fill='Maskpair') +
  facet_wrap(~year, scales = 'free_x', ncol=nrow(summary), switch="x") +
  theme(axis.text.x=element_text(angle = -60, hjust = 0))
        ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_grupper_v2.pdf"), width=16, height=9)


```

## Sorting by year and/or title

In statistics, we work with something called a *factor*. A factor is a variable or an observation which may hold values that cannot be sorted, as numerical data can. A color, for example, cannot be sorted - yet if you take them as factors, you can assign them a sequence you want them to have.

We might use factors to sort the plays by title. First, we can make a list of all titles sorted by year of publication. Then we can transform the title-column to factor variables, and assign the sequence that they had when they were sorted by year. 

```{r}
summary %>% 
  filter(speaker %in%   c("henrik", "jeronimus", "leander")) %>% 
  arrange(desc(year)) %>% # We sort the vertical axis by year (ascending)
  distinct(title) %>% 
  pull(title) -> factor_levels_sort_year

summary %>% 
  filter(speaker %in%   c("henrik", "jeronimus", "leander")) %>% 
  distinct(title) %>% 
  arrange(desc(title)) %>%  # We sort the vertical axis by title
  pull(title) -> factor_levels_sort_title

#summary$title_fct <- factor(summary$title, factor_levels)
```


## Sort by title

```{r}
summary %>% 
  #filter(speaker %in% c("henrik","pernille","troels")) %>%
  filter(speaker %in%   c("henrik", "jeronimus", "leander")) %>%
  arrange(title) %>%
  ggplot(aes(fill = speaker, y = freq, x = factor(title, factor_levels_sort_title))) +
    geom_bar(stat="identity") +
    theme(legend.position="bottom") +
    labs(
      title = "Distribution of spoken words between few characters that reoccur across Holberg's oeuvre by year of publication",
      #subtitle = toupper("TEST"),
      caption = "Source: DSL"
    ) +
  xlab("Play") +
  ylab("Percentage of spoken words") +
  facet_wrap(~year, ncol=nrow(summary), switch="x") +
  theme(axis.text.x=element_blank()) +
  coord_flip() # Flip the plot 90°
      ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_flip_v1.pdf"), width=16, height=9)

```


## Sort by year

```{r}
summary %>% 
  #filter(speaker %in% c("henrik","pernille","troels")) %>%
  filter(speaker %in%   c("henrik", "jeronimus", "leander")) %>%
  #arrange(year) %>%
  #mutate(title = str_c(title," ",year)) %>% 
  ggplot(aes(fill = speaker, y = 100*freq, x = factor(title, factor_levels_sort_year))) +
    geom_bar(stat="identity") +
    theme(legend.position="bottom") +
    labs(
      title = "Distribution of spoken words between few characters that reoccur across Holberg's oeuvre by year of publication",
      #subtitle = toupper("TEST"),
      caption = "Source: DSL"
    ) +
  xlab("Play") +
  ylab("Percentage of spoken words") +
  facet_wrap(~year, ncol=nrow(summary), switch="x") +
  #theme(axis.text.x=element_blank()) +
  coord_flip()
      ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_flip_v2.pdf"), width=16, height=9)

```

## Diagram that is not divided by year

First, we should make a new assignment of sequence to the factors, since we want to include years. 


```{r}
summary %>%
    ## filter(speaker %in%   c("henrik", "jeronimus", "leander")) %>% 
    mutate(title_fct2 = str_c(title," ", year)) -> summary_with_yeartitle

summary_with_yeartitle %>%
    arrange(desc(year)) %>% 
    distinct(title_fct2) %>% 
    pull(title_fct2) -> factor_levels_2

factor_levels_2

summary_with_yeartitle %>%
    mutate(title_fct2 = factor(title_fct2, factor_levels_2)) -> summary_fct2

summary_fct2 %>% select(title_fct2)
```

```{r}
summary_fct2 %>% 
    ## filter(speaker %in% c("henrik","pernille","troels")) %>%
    filter(speaker %in%   c("henrik", "jeronimus", "leander")) %>%
    ggplot(aes(fill = speaker, y = 100*freq, x = title_fct2)) +
       geom_bar(stat="identity") +
       theme(legend.position="bottom") +
       labs(
           title = "Distribution of spoken words between few characters that reoccur across Holberg's oeuvre by year of publication",
           ## subtitle = toupper("TEST"),
           caption = "Source: DSL"
       ) +
    scale_y_continuous(limits = c(0,100), breaks = c(0,50,100)) +
    xlab("Play") +
    ylab("Percentage of spoken words") +
    coord_flip()
      ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_flip_v3.pdf"), width=16, height=9)

```

```{r}
group_1 <- c("henrik", "jeronimus", "leander")
group_2 <- c("pernille", "leonora", "magdelone")

summary_fct2 %>%
    filter(speaker %in% c(group_1, group_2)) %>% 
    mutate(gruppe = if_else(speaker %in% group_1, "Henrik, Jeronimus og Leander","Pernille, Leonora og Magdelone")) %>% 
    ## filter(speaker %in%   c("henrik", "jeronimus", "leander")) %>%
    ggplot(aes(fill = gruppe, y = 100*freq, x = title_fct2)) +
       geom_bar(stat="identity") +
       theme(legend.position="bottom") +
       labs(
           title = "Distribution of spoken words between few characters that reoccur across Holberg's oeuvre by year of publication",
           ## subtitle = toupper("TEST"),
           caption = "Source: DSL"
       ) +
    scale_y_continuous(limits = c(0,100), breaks = c(0,10,20,30,40,50,60,70,80,90,100)) +
    xlab("Play") +
    ylab("Percentage of spoken words") +
      labs(fill='Maskpair') +
    coord_flip()
      ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_flip_grupper.pdf"), width=16, height=9)

```

## Let's Go Crazy

Below we show some examples of how to use different ggplot-themes. 

First we show the silly kind of example that nevertheless illustrates the wide range of thematic possibilities.

Before you can import and use the `tvthemes`, you need to install it first. You only need to do this once:


```{r}
#install.packages("tvthemes")
```

```{r}
library(tvthemes)
```

```{r}
summary %>% 
  #filter(speaker %in% c("henrik","pernille","troels")) %>%
  filter(speaker %in%   c("henrik", "jeronimus", "leander")) %>%
  #arrange(year) %>%
  ggplot(aes(fill = speaker, y = 100*freq, x = fct_rev(title))) +
    geom_bar(stat="identity") +
    theme_simpsons(
      #title.font = "Akbar",
      #text.font = "Akbar",
      axis.text.size = 8) +
    theme(legend.position="bottom") +
    labs(
      title = "Distribution of spoken words between characters",
      #subtitle = toupper("TEST"),
      caption = "Source: DSL"
    ) +
  xlab("Play") +
  ylab("Percentage of spoken words") +
  #facet_wrap(~year, ncol=nrow(summary), switch="x") +
  #theme(axis.text.x=element_blank()) +
  coord_flip()
      ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_tema_simpsons.pdf"), width=16, height=9)

```

## More usefu examples

In the following examples, we start by saving the plot to the variable `p`. After that, it is easy to plot first without sorting by year and thereafter sorting by year.


```{r}
summary %>% 
  #filter(speaker %in% c("henrik","pernille","troels")) %>%
  filter(speaker %in%   c("henrik", "jeronimus", "leander")) %>%
  arrange(year) %>%
  ggplot(aes(fill = speaker, y = 100*freq, x = factor(title, factor_levels_sort_title))) +
    geom_bar(stat="identity") +
    theme(legend.position="bottom") +
    labs(
      title = "Distribution of spoken words between characters",
      #subtitle = toupper("TEST"),
      caption = "Source: DSL"
    ) +
  xlab("Play") +
  ylab("Percentage of spoken words") +
  scale_fill_brewer(type = "seq", palette = "Greens") -> p # Choose a green palette

p + coord_flip() +
  facet_wrap(~year, ncol=nrow(summary), strip.position="top") 
      ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_greens.pdf"), width=16, height=9)

```

## Viridis

Another package. Again, we only need to install this package once the first time we run this notebook. After that, we can simply import it.

```{r}
#install.packages("viridis")
```

```{r}
library(viridis)
```

```{r}
summary %>% 
  #filter(speaker %in% c("henrik","pernille","troels")) %>%
  filter(speaker %in%   c("henrik", "jeronimus", "leander")) %>%
  arrange(year) %>%
  ggplot(aes(fill = speaker, y = 100*freq, x = factor(title, factor_levels_sort_year))) +
    geom_bar(stat="identity") +
    theme(legend.position="bottom") +
    labs(
      title = "Distribution of spoken words between characters",
      #subtitle = toupper("TEST"),
      caption = "Source: DSL"
    ) +
  xlab("Play") +
  ylab("Percentage of spoken words") +
  scale_fill_viridis(discrete = TRUE) +
  coord_flip() -> p

p
      ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_viridis_v1.pdf"), width=16, height=9)

p + facet_wrap(~year, ncol=nrow(summary), strip.position="top")
      ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_viridis_v2.pdf"), width=16, height=9)

```

## Economist

Another. Again, install `ggthemes` only the first time you run this notebook.

```{r}
#install.packages("ggthemes")
```

```{r}
library(ggthemes)
```

```{r}
summary %>% 
  #filter(speaker %in% c("henrik","pernille","troels")) %>%
  filter(speaker %in%   c("henrik", "jeronimus", "leander")) %>%
  arrange(year) %>%
  ggplot(aes(fill = speaker, y = 100*freq, x = factor(title, factor_levels_sort_year))) +
    geom_bar(stat="identity") +
    theme(legend.position="bottom") +
    labs(
      title = "Distribution of spoken words between characters",
      #subtitle = toupper("TEST"),
      caption = "Source: DSL"
    ) +
  xlab("Play") +
  ylab("Percentage of spoken words") +
  #scale_fill_brewer(palette = "Greens") +
  theme_economist() + scale_fill_economist() +
  coord_flip() -> p

p
      ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_tema_economist_v1.pdf"), width=16, height=9)

p + facet_wrap(~year, ncol=nrow(summary), strip.position="top")
      ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_tema_economist_v2.pdf"), width=16, height=9)

```

## Minimal

```{r}
summary %>% 
  #filter(speaker %in% c("henrik","pernille","troels")) %>%
  filter(speaker %in%   c("henrik", "jeronimus", "leander")) %>%
  arrange(year) %>%
  ggplot(aes(fill = speaker, y = 100*freq, x = factor(title, factor_levels_sort_year))) +
    geom_bar(stat="identity") +
    theme(legend.position="bottom") +
    labs(
      title = "Distribution of spoken words between characters",
      #subtitle = toupper("TEST"),
      caption = "Source: DSL"
    ) +
  xlab("Play") +
  ylab("Percentage of spoken words") +
  #scale_fill_brewer(palette = "Greens") +
  facet_wrap(~year, ncol=nrow(summary), strip.position = "top") +
  #theme(axis.text.x=element_blank()) +
  coord_flip() +
  #theme_bw()
  theme_minimal()
      ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_tema_minimal.pdf"), width=16, height=9)

```

## Let's do the comparison between groups once more

```{r}
group_1 <- c("henrik", "jeronimus", "leander")
group_1_name <- "Henrik, Jeronimus og Leander"

group_2 <- c("pernille", "leonora", "magdelone")
group_2_name <- "Pernille, Leonora og Magdelone"

summary %>% 
  filter(!is.na(title)) %>% 
  filter(speaker %in% c(group_1, group_2)) %>% 
  mutate(gruppe = if_else(speaker %in% group_1, group_1_name,group_2_name)) %>% 
  ggplot(aes(fill = gruppe, y = freq*100, x = fct_rev(title))) +
    geom_bar(stat="identity") +
    theme(legend.position="bottom") +
    labs(
      title = "Distribution of spoken words between few characters that reoccur across Holberg's oeuvre by year of publication",
      #subtitle = toupper("TEST"),
      caption = "Source: DSL"
    ) +
  xlab("Play") +
  ylab("Percentage of spoken words") +
  labs(fill='Maskpair') +
  coord_flip()
      ggsave(here("graphs/pers-visualiseringer", "fordeling_talte_ord_grupper_igen.pdf"), width=16, height=9)

```

# Coloring

Before running this section, we need to have previously computed `summary_fct2`.

```{r}
summary_fct2 %>% 
    ## filter(speaker %in% c("henrik","pernille","troels")) %>%
    filter(speaker %in%   c("henrik", "jeronimus", "leander")) %>%
    ggplot(aes(fill = speaker, y = 100*freq, x = title_fct2)) +
       geom_bar(stat="identity") +
       theme(legend.position="bottom") +
       labs(
           title = "Distribution of spoken words between few characters that reoccur across Holberg's oeuvre by year of publication",
           ## subtitle = toupper("TEST"),
           caption = "Source: DSL"
       ) +
    scale_y_continuous(limits = c(0,100), breaks = c(0,50,100)) +
    xlab("Play") +
    ylab("Percentage of spoken words") +
    coord_flip()
```

Now we want to assign our own colors to Henrik, Jeronimus and Leander.


```{r}
summary_fct2 %>%
    filter(speaker %in%   c("henrik", "jeronimus", "leander")) %>%
    ggplot(aes(fill = speaker, y = 100*freq, x = title_fct2)) +
       geom_bar(stat="identity") +
       theme(legend.position="bottom") +
       labs(
           title = "Distribution of spoken words between few characters that reoccur across Holberg's oeuvre by year of publication",
           ## subtitle = toupper("TEST"),
           caption = "Source: DSL"
       ) +
      scale_fill_manual( # Here we fill in the colors of characters manually
        values = c(
          "henrik" = "purple",
          "jeronimus" = "orange",
          "leander"="steelblue"
        )
      ) +
    scale_y_continuous(limits = c(0,100), breaks = c(0,50,100)) +
    xlab("Play") +
    ylab("Percentage of spoken words") +
    coord_flip()
```

# Finally, save a visualization as a horizontal A4 pdf-file:

```{r}
ggsave(here("graphs/pers-visualiseringer", "ng3.pdf"),
       width = 297, 
       height = 210, 
       units = "mm")
```
