---
title: "Plots"
output: html_notebook
author: "baj"
---

# Plot who says what when

This notebook will simply show another way of visualizing when and what characters speak in a play. 

# Import libraries and read in the data

```{r, message=FALSE}
library(tidyverse)
library(tidytext)
library(readxl)
library(here)
library(ndjson)
#library(xslt)
library(fs)
library(ggplot2)
library(knitr)
library(dplyr)

require(readr)  # We will need this to read csv files (with "read_csv()")
require(purrr)  # We will need this to use the functions "map()" and "reduce()"

source(here("src/p017-functions.R"))
source(here("src", "present_without_speech.R"))

#convert_TEI_to_JSONL(here("test-data"))
# my_file <- "Skibbrud_mod.jsonl"
# my_file <- "Sganarel_mod.jsonl"
my_file <- "Barselstuen_mod.jsonl" 
# my_file <- "Mascarade_mod.jsonl"
my_play <- read_play_jsonl(here("test-data", my_file))
my_play
```


## Variations of character-names

```{r}

# Read in the excel-sheet of names of characters and name-variations in the plays
variants <- read_excel(here("Rolleliste.xlsx")) %>% 
  unnest_tokens(variant, Alias, token = "regex", pattern = ", ") %>% # Collect all the names and name-variations from the excel-sheet.
    mutate(
      Karakter = tolower(Karakter),
      variant = tolower(variant)
    )
# Use the excel-sheet to collect all variations of character-names in the play
my_play %>%
  mutate(speaker = tolower(speaker)) %>% 
  left_join(variants, by = c("filename"="Filnavn", "speaker"="variant")) %>% 
  mutate(speaker = if_else(!is.na(Karakter), Karakter, speaker)) %>%
  filter(!is.na(speaker), !(speaker=="")) %>%
  distinct -> my_play
```

## Add an index for scene

```{r, warning=FALSE, message=FALSE}
my_play %>% 
   group_by(act_number)  %>% 
   summarise(scenes = max(scene_number)) %>% 
   pull(scenes) %>% cumsum() -> act_length

act_length

calc_scene_index <- function(act_number, scene_number, act_length) {
   ifelse(is.numeric(scene_number),
      scene_number + ifelse(act_number==1, 0, act_length[act_number-1]),
      NA
   )
}

```

## Restructure the data

It will look nicer if we write it into the csv. 
Remember to count how many characters are present.


```{r, rows.print=20}
my_play %>% 
  # Add a scene index for the x-axis
  rowwise() %>% 
  mutate(scene_index = calc_scene_index(act_number, scene_number, act_length)) %>% 
  # Remove rows that are not dialogue
  filter(act != "", scene != "", speaker != "", !is.na(spoke)) %>%
  # Make a boolean to indicate spoke
  mutate(boolean_spoke = (if (spoke != "") {"X"} else {"Y"})) %>%
  # Keep only coloumns that we need
  select(act_number, scene_number, scene_index, speaker, boolean_spoke) %>%
  distinct() -> my_summary

(my_summary %>%
  # Make a small scene_index, act_number, scene_number table
  select(scene_index, act_number, scene_number) %>% 
  distinct() -> my_acts_and_scenes)
(my_acts_and_scenes %>% pull(act_number) -> my_acts)



my_summary %>%
  # Select only the columns that we are interested in
  select(scene_index, speaker, boolean_spoke) %>% 
  distinct() %>%
  # Now pivot speakers to rows
  pivot_wider(names_from = scene_index, values_from = boolean_spoke) -> my_speakers

# Now remove NA
my_speakers[is.na(my_speakers)] <- ""
my_speakers

# Write it to a new csv
write.csv(my_speakers, file = here(paste("csv/","plot_who_speaks_", my_file, ".csv")))

# Pretty print my_speakers
# install.packages("DT") # Install DT first (only the first time you use it)
library(DT)
datatable(my_speakers)

```


# Plotting who speaks

```{r}
my_play %>% 
  select(docTitle) %>% 
  distinct() %>%
  pull(1,1) -> my_title

my_play %>% 
  select(year) %>% 
  distinct() %>%
  pull(1,1) -> my_year

# Remember to count how many characters are on stage now
(my_summary %>%
  group_by(act_number, scene_number) %>%
  mutate(total = sum(n_distinct(speaker))) -> my_summary) 

# Let's plot it
my_summary %>% 
  ggplot(aes(y = speaker, x = scene_index, width = 100)) +
    geom_tile(aes(fill = boolean_spoke, colour = speaker), colour = "grey", show.legend = FALSE) + 
    geom_text(aes(label = "X")) +
    labs(
      title = paste(my_title, my_year, my_file),
      subtitle = "Who speaks when"
    ) +
  xlab("Act \nScene \nNumber of characters speaking in the scene") +
  ylab("Speaker") + 
  theme(axis.text.x=element_blank(), line = element_blank(), rect = element_blank()) + 
  facet_grid(cols = vars("act_number" = act_number, "scene_number" = scene_number, "total" = total), switch="x")

# And then we save the plot
  ggsave(here("graphs/plots", paste(my_file, ".hvem_taler_hvornaar.pdf")), width=16, height=9)
  

```

# Plotting who is present

```{r}
# First, let's read in the play again
my_play <- read_play_jsonl(here("test-data", my_file))
# And find out who is present but does not speak
present_without_speech_numbers(my_play) %>%
  # Add a scene index for the x-axis
  rowwise() %>% 
  mutate(scene_index = calc_scene_index(act_number, scene_number, act_length)) %>%
  # Add a boolean to denote who is present
    mutate(boolean_spoke = "S") %>%
  select(-index) -> present_silent 

# Combine those who speak with those are present
my_summary %>%
    select(-total) %>%
  full_join(present_silent, 
            by = c("speaker"="word", "act_number", "scene_number", "scene_index", "boolean_spoke")) -> new_summary

# Remember to count how many characters are on stage now
(new_summary %>%
  group_by(act_number, scene_number) %>%
  mutate(total = sum(n_distinct(speaker))) -> new_summary) 

new_summary %>% 
  ggplot(aes(y = speaker, x = scene_index, width = 100)) +
    geom_tile(aes(fill = boolean_spoke, colour = speaker), colour = "grey", show.legend = FALSE) + 
    geom_text(aes(label = boolean_spoke)) +
    labs(
      title = paste(my_title, my_year, my_file),
      subtitle = "Who is present"
    ) +
  xlab("Act \nScene \nNumber of characters in the scene") +
  ylab("Speaker") + 
  theme(axis.text.x=element_blank(), line = element_blank(), rect = element_blank()) + 
  facet_grid(cols = vars("act_number" = act_number, "scene_number" = scene_number, "total" = total), switch = "x")

  #ggsave(here("graphs/plots", paste(my_file, ".hvem_til_stede.pdf")), width=16, height=9)

```

# Plotting who is spoken of

We want to know which characters are spoken of, even if they themselves are not present in the scene.
First we need to define what we mean by being "spoken about". We might find characters who are designated as "not present", while being mentioned in the speech of another character. So if a character is mentioned while not being present in the scene, the character is "spoken about".

Below we will find these characters and can add this data to the above plot.

```{r}
source(here("src", "omtale.R"))
(spoken_about(my_play) %>%
  # Add a scene index for the x-axis
  rowwise() %>% 
  mutate(scene_index = calc_scene_index(act_number, scene_number, act_length)) %>%
  # Add a boolean for "spoken about"
    mutate(boolean_spoke = "O") %>%
  select(-index)
  -> spoken_about_summary) 

new_summary %>%
    select(-total) %>%
  full_join(spoken_about_summary, by = c("speaker"="word", "act_number", "scene_number", "scene_index", "boolean_spoke")) %>% 
  ggplot(aes(y = speaker, x = scene_index, width = 100)) +
    geom_tile(aes(fill = boolean_spoke, colour = speaker), colour = "grey", show.legend = FALSE) + 
    geom_text(aes(label = boolean_spoke)) +
    labs(
      title = paste(my_title, my_year, my_file),
      subtitle = "Who is present and who is spoken about"
    ) +
  xlab("Act \nScene") +
  ylab("Speaker") + 
  theme(axis.text.x=element_blank(), line = element_blank(), rect = element_blank()) + 
  facet_grid(cols = vars("act_number" = act_number, "scene_number" = scene_number), switch = "x")

  #ggsave(here("graphs/plots", paste(my_file, ".hvem_til_stede_hvem_omtalt.pdf")), width=16, height=9)

```

# Let's apply these functions to all plays

Let's first test this on a single play.

```{r}
source(here("src", "plot_func.R"))

my_file <- "Mascarade_mod.jsonl"
plot_all(my_file)

```

And now on all plays

```{r}
source(here("src", "plot_func.R"))
# Get all files in the JSONL format
json_file_list <- list.files(here("test-data"), pattern = "*.jsonl")
# Get plots for all plays
for(i in 2:length(json_file_list)) {
  plot_all(json_file_list[i])
  print(paste("plotting play nr.",i+1,"of",length(json_file_list)))
}

```

