---
title: "Who is present without speaking?"
output: html_notebook
author: baj
---

# Defining who is present but does not speak

We can find out who is present without speaking by using the stage-directions. 

We import the necessary R packages before we start.


```{r, message=FALSE}
library(tidyverse)
library(ndjson)
library(here)
library(fs)
#library(xslt)
library(readxl)
library(tidytext)
```

We also need to load the functions created for this project. This is done by `source`-ing them.

```{r}
source(here("src", "p017-functions.R"))
source(here("src", "present_without_speech.R"))
```

# Read play

(Look at read-plays.Rmd for help)

In this example I use "Barselstuen_mod_nc.page"
You can choose any play you would like to. Create a "test-data" folder in your project directory, and copy the play here.
Edit the file below to the JSONL-file of the play you want. Simply load it into the 'my_file' variable

```{r}
my_file <- "Mascarade_mod.jsonl"
# Convert all page files in the specified folder to jsonl
# convert_TEI_to_JSONL(here("test-data")) 
# Read the specified play from the jsonl file
play <- read_play_jsonl(here("test-data", my_file))
play # Show the table for the play

# Remember to load in the excel sheet of variations of the names of characters.
# Read in the excel sheet
variants <- read_excel(here("Rolleliste.xlsx")) %>% 
  unnest_tokens(variant, Alias, token = "regex", pattern = ", ") %>% 
    mutate(
      Karakter = tolower(Karakter),
      variant = tolower(variant)
    )
# Use the excel sheet to gather and fuse variations of character-names
play %>%
  mutate(speaker = tolower(speaker)) %>% 
  left_join(variants, by = c("filename"="Filnavn", "speaker"="variant")) %>% 
  mutate(speaker = if_else(!is.na(Karakter), Karakter, speaker)) %>%
  filter(!is.na(speaker), !(speaker=="")) %>%
  distinct -> play
```

The variable 'play' is now the play in question. This way, you can just change that and rerun all the sections below, without having to change them.

# Find speakers

Find all designated speakers

```{r}
(play %>%
  filter(!is.na(speaker)) %>% 
  count(speaker) %>%
  # mutate(speaker = str_to_lower(speaker)) %>%
  select(speaker) -> speakers)
```

# Find all stage directions

Stage directions exist on two levels, preceded by either <stage> or by <speaker_stage> in the JSONL-files.

```{r}

# These are the people who are directly mentioned in the <stage> tokens
(play %>% 
  filter(!is.na(stage)) %>%
  filter(!startsWith(stage, "("))  %>% 
  unnest_tokens(word, stage, drop=FALSE, token="regex", pattern = ", *") %>% # Tokenize <stage>
  select(act, scene, index, word) %>% 
  distinct() -> explicit_stage_tokens)
  
# These are the the actors who are implicitly mentioned in <stage> tokens
(play %>% 
  filter(!is.na(stage)) %>%
  filter(startsWith(stage, "("))  %>% 
  unnest_tokens(word, stage) %>% # Tokenize <stage>
  select(act, scene, index, word) %>% 
  distinct() -> implicit_stage_tokens)
  
# These are the the actors who are implicitly mentioned in <speaker_stage> tokens
(play %>% 
  unnest_tokens(word, speaker_stage) %>% # Tokenize <speaker_stage>
  filter(!is.na(word)) %>%
  select(act, scene, index, word) -> speaker_stage_tokens)
```

# Search for speakers in the stage directions

```{r}
(explicit_stage_tokens %>%
  semi_join(speakers, by = c("word" = "speaker")) -> explicit_speakers_in_stage)

(implicit_stage_tokens %>%
  semi_join(speakers, by = c("word" = "speaker")) -> implicit_speakers_in_stage)

(speaker_stage_tokens %>%
  semi_join(speakers, by = c("word" = "speaker")) -> speakers_in_speaker_stage)


(explicit_speakers_in_stage %>%
    full_join(implicit_speakers_in_stage) %>% 
    full_join(speakers_in_speaker_stage) -> all_speakers_in_stage)

```

The output we get here must mean that we are finding 101 places where a speaker is mentioned in a line of stage directions. That is, considering only the list of all speakers that we know are in the play.

# Remove characters that are actually speaking

```{r}
         
## Distinct speakers in each scene in each act   
(play %>% 
  filter(!is.na(speaker)) %>%
  select(act, scene, speaker) %>%
  mutate(speaker = str_to_lower(speaker)) %>%
  distinct() -> distinct_speakers)

## Filter out speakers from words grouped by act and scene
all_speakers_in_stage %>%
  anti_join(distinct_speakers, by=c("act"="act", "scene"="scene", "word"="speaker")) %>% 
  distinct()

```

Now we see that there are 18 instances where a character is mentioned in the stage directions while that character is not speaking in the scene.

But is it reliable?

In act 2, scene 1, we find: <speaker>En pige <stage>(med en tallerken)</stage></speaker>. However, our algorithm cannot detect that "pigen" in <stage> is exactly the same as "En pige" in <speaker>, who is speaking... We have an error because it is counting the girl twice.

We have tested this on "Barselstuen", since we know that in this play, we need to find Corfitz lying under the bed (present but not speaking) in act 2, scene 2 - and we can! However, it is necessary go through the findings manually so as to sort out instances of the sort of "en pige" and "pigen", mentioned before.

If we want to use this in the plot, we should make a function of all of it.

We might then also try the function on another play, and maybe get a different result?

# Let's test the function

```{r}
# Read the play in again to make it work
play <- read_play_jsonl(here("test-data", my_file))
present_without_speech(play)
```

The function seems to work, so we can add it to the visualization of the network. 

