---
title: "Visualizing percentages of spoken words across scenes and plays"
output: html_notebook
author: db
---

Loading required libraries and functions.

```{r, message=FALSE}
library(tidyverse)
library(tidytext)
library(readxl)
library(here)
library(ndjson)
#library(xslt)
library(fs)
library(ggplot2)

require(readr)  # For read_csv()
require(purrr)  # For map(), reduce()

source(here("src/p017-functions.R"))
```

One of the P017 functions is used to load all plays by Holberg, located in the `test-data`-folder, into a table called allplays. Then, characters, that are referred to with their alias, are renamed to their main name.

```{r}
allplays <- read_plays_jsonl(here("test-data")) %>% 
  # Add column containing both tidy title and year
  mutate(title_year = paste0(title, " ", year))

# Load excel sheet with name variants
variants <- read_excel(here("Rolleliste.xlsx")) %>% 
  unnest_tokens(variant, Alias, token = "regex", pattern = ", ") %>% 
  mutate(
      Karakter = tolower(Karakter),
      variant = tolower(variant))

# Use the variants object to rename certain speakers in allplays
allplays <- allplays %>%
  mutate(speaker = tolower(speaker)) %>% 
  left_join(variants, by = c("filename"="Filnavn", "speaker"="variant")) %>% 
  mutate(speaker = if_else(!is.na(Karakter), Karakter, speaker))

```

Groups of interest are set and colors are assigned. You can find the hex code of colors by searching on "color picker" in Google.

```{r}
group1 <- c("henrik", "jens")
color1 <- "#2CB6DF" 
group2 <- c("barselskvinden")
color2 <- "#CF4426"
group3 <- c("jacob")
color3 <- "#672e99"
group4 <- c()
color4 <- NA
group5 <- c()
color5 <- NA
group6 <- c()
color6 <- NA
group7 <- c()
color7 <- NA
group8 <- c()
color8 <- NA
group9 <- c()
color9 <- NA
group10 <- c()
color10 <- NA
group11 <- c()
color11 <- NA
group12 <- c()
color12 <- NA


# Combine the speakers into a vector
my_speakers <- c(group1, group2, group3, group4, group5, group6, group7, group8, group9, group10, group11, group12)

my_colors <- c(color1, color2, color3, color4, color5, color6, color6, color8, color9, color10, color11, color12) %>% na.omit()

my_plays <- unique(allplays$docTitle[allplays$speaker %in% my_speakers])
```

A table is prepared for plotting.

```{r}
df <- allplays %>% 
  
  # Remove rows that are not dialogue
  filter(act != "", scene != "", speaker != "", !is.na(speaker), !is.na(spoke)) %>%
  
  # Keep only plays of interest
  filter(docTitle %in% my_plays) %>% 
  
  # Add the number of spoken words
  mutate(n_spoken_words = str_count(spoke, '\\w+')) %>% 
  
  # Organize dataset by grouping
  group_by(year, title_year, act_number, scene_number, speaker) %>% 
  
  # Sum the words spoken by each speaker
  summarise(words = sum(n_spoken_words))

# Find highest number of scenes within each act for use in the plot  
intercepts <- df %>% group_by(act_number) %>% 
  summarise(max_scene = max(scene_number)) %>% 
  mutate(intercepts = cumsum(max_scene+c(0.5,rep(0,length(act_number)-1)))) %>%
  pull(intercepts)

# Make order of plays chronological   
play_chronology <- unique(df$title_year)
df$title_year <- factor(df$title_year, levels = play_chronology)
  
df <- df %>%   
  # Calculate percentage of words spoken in each scene
  group_by(title_year, act_number, scene_number) %>% 
  mutate(percent = 100*(words/sum(words))) %>% ungroup() %>% 

  # Add act:scene column
  mutate(act_scene = paste0(act_number, ":", str_pad(scene_number, 2, pad = "0")))

# Make group names into text that can be showed in the plot legend
for (i in 1:12){
  text <- paste(eval(parse(text = paste0("group", i))), collapse = ", ")
  assign(paste0("group", i, "text"), text)
}

# Make a column that indicates group
df$group <- "other"

for (i in 1:12){
  df$group <- ifelse(df$speaker %in% eval(parse(text = paste0("group", i))) & df$group=="other", eval(parse(text = paste0("group", i, "text"))), df$group)
}

groups_used <- c(group12text, group11text, group10text, group9text, group8text, group7text, group6text, group5text, group4text, group3text, group2text, group1text) %>% as_tibble() %>% filter(value != "") %>% pull(value)
  
df$group <- factor(df$group, levels = c("other", groups_used))
  
# If two characters from the same group speak in the same scene, their percentages should be added
df <- df %>% group_by(title_year, act_scene, group) %>% 
   summarise(percent_sum = sum(percent))

```

```{r}
  ggplot(df, aes(fill = group, y = percent_sum, x = act_scene)) +
    geom_bar(stat="identity") +
  scale_fill_manual(breaks = c(groups_used[length(groups_used):1], "other"), 
                    values = c(my_colors, "grey")) +
  xlab("Act:Scene") +
  ylab("Percentage of spoken words") +
  facet_grid(rows = vars("title_year" = title_year), 
             switch = "x") +
  scale_y_continuous(breaks = c(0, 50, 100)) +
  geom_vline(xintercept = intercepts, size = 0.2) +
  theme_bw() +
  theme(legend.position = "bottom", 
          legend.title = element_blank(), 
          axis.text.x = element_text(angle = 90, size = 6, hjust = 0, vjust = 0.5),
          axis.text.y = element_text(size = 4),
          strip.text.y = element_text(angle = 0, hjust = 0))
```


Make a plot like the one above, but now it shows percentage of words spoken in the whole play and not just one particular scene. A df is prepared again, and the plot is drawn.

```{r}
df <- allplays %>% 
  
  # Remove rows that are not dialogue
  filter(act != "", scene != "", speaker != "", !is.na(speaker), !is.na(spoke)) %>%
  
  # Keep only plays of interest
  filter(docTitle %in% my_plays) %>% 
  
  # Add the number of spoken words
  mutate(n_spoken_words = str_count(spoke, '\\w+')) %>% 
  
  # Organize dataset by grouping
  group_by(year, title_year, act_number, scene_number, speaker) %>% 
  
  # Sum the words spoken by each speaker
  summarise(words = sum(n_spoken_words))

# Find highest number of scenes within each act for use in the plot  
intercepts <- df %>% group_by(act_number) %>% 
  summarise(max_scene = max(scene_number)) %>% 
  mutate(intercepts = cumsum(max_scene+c(0.5,rep(0,length(act_number)-1)))) %>%
  pull(intercepts)

# Make order of plays chronological   
play_chronology <- unique(df$title_year)
df$title_year <- factor(df$title_year, levels = play_chronology)
  
df <- df %>%   
  # Calculate percentage of words spoken in each scene
  group_by(title_year) %>% mutate(percent = 100*(words/sum(words))) %>% ungroup() %>% 

  # Add act:scene column
  mutate(act_scene = paste0(act_number, ":", str_pad(scene_number, 2, pad = "0")))

# Make group names into text that can be showed in the plot legend
for (i in 1:12){
  text <- paste(eval(parse(text = paste0("group", i))), collapse = ", ")
  assign(paste0("group", i, "text"), text)
}

# Make a column that indicates group
df$group <- "other"

for (i in 1:12){
  df$group <- ifelse(df$speaker %in% eval(parse(text = paste0("group", i))) & df$group=="other", eval(parse(text = paste0("group", i, "text"))), df$group)
}

groups_used <- c(group12text, group11text, group10text, group9text, group8text, group7text, group6text, group5text, group4text, group3text, group2text, group1text) %>% as_tibble() %>% filter(value != "") %>% pull(value)
  
df$group <- factor(df$group, levels = c("other", groups_used))
  
# If two characters from the same group speak in the same scene, their percentages should be added
df <- df %>% group_by(title_year, act_scene, group) %>% 
   summarise(percent_sum = sum(percent))

# PLOTTING
ggplot(df, aes(fill = group, y = percent_sum, x = act_scene)) +
  geom_bar(stat="identity") +
  scale_fill_manual(breaks = c(groups_used[length(groups_used):1], "other"), 
                    values = c(my_colors, "grey")) +
  xlab("Act:Scene") +
  ylab("Percentage of spoken words") +
  facet_grid(rows = vars("title_year" = title_year), 
             switch = "x", scales = "free_y") +
  #scale_y_continuous(breaks = c(0, 50, 100)) +
  geom_vline(xintercept = intercepts, size = 0.2) +
  theme_bw() +
  theme(legend.position = "bottom", 
          legend.title = element_blank(), 
          axis.text.x = element_text(angle = 90, size = 6, hjust = 0, vjust = 0.5),
          axis.text.y = element_text(size = 4),
          strip.text.y = element_text(angle = 0, hjust = 0))

```

Make a plot like the one above but now, the group "other" differentiates between social classes. This requires the social class data to be combined with allplays before hand.

```{r}
# Load social class data
social <- read_excel(here("gender AND Mask_alaw sammenlagte karakterer reduceret i kategoriantal.xlsx"))
social <- unite(data = social, col = "social_status", "social status, main character", "social status, other characters", sep="", na.rm = TRUE) %>% 
  mutate(social_status = gsub("^NA", "", social_status)) %>% 
  mutate(social_status = gsub("NA$", "", social_status))

# Combine with allplays
allplays <- allplays %>%
  mutate(docTitle = tolower(docTitle)) %>% 
  left_join(social[,c(1,2,7)], by = c("docTitle"="play", "speaker"="speaker"))

# Decapitalize my_plays
my_plays <- unique(allplays$docTitle[allplays$speaker %in% my_speakers])

# PREPARE THE TABLE
df <- allplays %>% 
  
  # Remove rows that are not dialogue
  filter(act != "", scene != "", speaker != "", !is.na(speaker), !is.na(spoke)) %>%
  
  # Keep only plays of interest
  filter(docTitle %in% my_plays) %>% 
  
  # Add the number of spoken words
  mutate(n_spoken_words = str_count(spoke, '\\w+')) %>% 
  
  # Organize dataset by grouping
  group_by(year, title_year, act_number, scene_number, speaker, social_status) %>% 
  
  # Sum the words spoken by each speaker
  summarise(words = sum(n_spoken_words))

# Find highest number of scenes within each act for use in the plot  
intercepts <- df %>% group_by(act_number) %>% 
  summarise(max_scene = max(scene_number)) %>% 
  mutate(intercepts = cumsum(max_scene+c(0.5,rep(0,length(act_number)-1)))) %>%
  pull(intercepts)

# Make order of plays chronological   
play_chronology <- unique(df$title_year)
df$title_year <- factor(df$title_year, levels = play_chronology)
  
df <- df %>%   
  # Calculate percentage of words spoken in each scene
  group_by(title_year) %>% mutate(percent = 100*(words/sum(words))) %>% ungroup() %>% 

  # Add act:scene column
  mutate(act_scene = paste0(act_number, ":", str_pad(scene_number, 2, pad = "0")))

# Make group names into text that can be showed in the plot legend
for (i in 1:12){
  text <- paste(eval(parse(text = paste0("group", i))), collapse = ", ")
  assign(paste0("group", i, "text"), text)
}

# Make a column that indicates group
df$group <- ifelse(df$social_status=="", "Unknown", df$social_status)
statuses <- unique(df$social_status)

for (i in 1:12){
  df$group <- ifelse(df$speaker %in% eval(parse(text = paste0("group", i))) & df$social_status %in% statuses, eval(parse(text = paste0("group", i, "text"))), df$group)
}

groups_used <- c(group12text, group11text, group10text, group9text, group8text, group7text, group6text, group5text, group4text, group3text, group2text, group1text) %>% as_tibble() %>% filter(value != "") %>% pull(value)

statuses_used <- unique(df$group[!df$group %in% groups_used])
  
df$group <- factor(df$group, levels = c(statuses_used, groups_used))
  
# If two characters from the same group speak in the same scene, their percentages should be added
df <- df %>% group_by(title_year, act_scene, group) %>% 
   summarise(percent_sum = sum(percent))

# PLOTTING
library(colorspace)

ggplot(df, aes(fill = group, y = percent_sum, x = act_scene)) +
  geom_bar(stat="identity") +
  scale_fill_manual(breaks = c(groups_used[length(groups_used):1], statuses_used), 
                    values = c(my_colors, qualitative_hcl(length(statuses_used), palette = "Set 3"))) +
  xlab("Act:Scene") +
  ylab("Percentage of spoken words") +
  facet_grid(rows = vars("title_year" = title_year), 
             switch = "x", scales = "free_y") +
  #scale_y_continuous(breaks = c(0, 50, 100)) +
  geom_vline(xintercept = intercepts, size = 0.2) +
  theme_bw() +
  theme(legend.position = "bottom", 
          legend.title = element_blank(), 
          axis.text.x = element_text(angle = 90, size = 6, hjust = 0, vjust = 0.5),
          axis.text.y = element_text(size = 4),
          strip.text.y = element_text(angle = 0, hjust = 0))

```

The code making the two plots above has been gathered in two different functions.

```{r}
source(here("src", "percent-speech-functions.R"))

percentage_plot(group1 = "leander", group2 = c("pernille", "barselskvinden"), color1 = "#2CB6DF", color2 = "#CF4426")

percentage_plot_status(group1 = "leander", group2 = c("pernille", "barselskvinden"), color1 = "#2CB6DF", color2 = "#CF4426")
```

