---
title: "comparing-auto-and-manual'"
author: "TAJE"
date: "2021-09-24"
output: html_document
---

Loading libraries
```{r}
library(tidyverse)
library(readxl)
library(here)
library(fs)
source(here("src", "p017-functions.R"))
```

Load the manual excelsheet
```{r}
excel_sheet <- read_excel(here("allspeakers_gender_and_mask.xlsx"))
```

```{r}
excel_sheet %>% 
  select(-`...1`) %>% # hvor kommer denne kolonne fra?
  separate_rows(plays,sep = ",") %>% 
  rename(play = plays) -> manual
```

Load the auto

```{r}
read_plays_jsonl(here("test-data")) -> allplays
```

```{r}
allplays %>% 
  filter(!is.na(speaker) & speaker != "") %>% 
  count(speaker,docTitle) %>% 
  rename(play = docTitle) -> auto
```

Create new tables with all speakers and play titles down cased

```{r}
allplays %>% 
  filter(!is.na(speaker) & speaker != "") %>%
  rename(play = docTitle) %>% 
  mutate(
    speaker = str_to_lower(speaker),
    play = str_to_lower(play)
  ) %>% 
  count(speaker,play) -> auto_decapitalized

manual %>% 
  mutate(
    speaker = str_to_lower(speaker),
    play = str_to_lower(play)
  ) -> manual_decapitalized
```

Delete spaces:
```{r}
manual_decapitalized %>% 
  mutate(play = str_trim(play)) -> manual_decapitalized_trimmed
```

So which combinations of speaker and play are only in manual and not in auto?
```{r}
manual_decapitalized_trimmed %>%
  select(speaker, play) %>%
  setdiff(auto_decapitalized %>% select(speaker, play))
```
So 85 combinations that are not in auto.

And which combinations of speaker and play are only in manual and not in auto?
```{r}
auto_decapitalized %>% 
  select(speaker, play) %>% 
  setdiff(manual_decapitalized_trimmed %>% select(speaker, play))
```
142 combinations that are not in manual.

Try to combine the two tables

```{r}
auto_decapitalized %>% 
  mutate(in_auto = TRUE) %>% 
  full_join(
    manual_decapitalized_trimmed %>%
      mutate(in_manual = TRUE),
    by = c("speaker","play")
  ) %>% 
  mutate(
    in_auto = !is.na(in_auto),
    in_manual = !is.na(in_manual)
  ) %>% 
  select(speaker, play, in_auto, in_manual, everything()) -> gender_and_mask
```

```{r}
write_excel_csv2(gender_and_mask,here("gender-and-mask.csv"))
```

